---
title: "Exercises for class 1"
output: html_document
date: "13-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
```

# Exercises for Class 1

## Supporting code

Below is code from Chapter 2 of the book that you can use to do the exercises.
Not all the code you require is here though.
Make sure to not rush it and take time to play around with each function.
:))

```{r}
# Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20)

# compute likelihood at each value in grid
likelihood = dbinom(6, size = 9, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

## Exercises

### Easy

2E1.

```{r}

(2) Pr(rain|Monday)
# (4) Pr(rain, Monday)/ Pr(Monday

```

2E2.

```{r}

(3) The probability that it is Monday, given that it is raining

```

2E3.

```{r}

(1) Pr(Monday|rain)
# (4) Pr(rain|Monday) Pr(Monday)/ Pr(rain)

```

2E4.

```{r}

The probability of water being 0.7 reflects our subjective estimate based on available data, not an exact truth. Therfore it is subject to change as the new data is presented, and not a firm statement.

```

### Medium

2M1.

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 100)
prior <- rep(1, 100)

# (1) W, W, W
likelihood_1 <- dbinom(3, size = 3, prob = p_grid)
posterior_1 <- likelihood_1 * prior / sum(likelihood_1 * prior)
plot(posterior_1 ~ p_grid, type = "l")

# (2) W, W, W, L
likelihood_2 <- dbinom(3, size = 4, prob = p_grid)
posterior_2 <- likelihood_2 * prior / sum(likelihood_2 * prior)
plot(posterior_2 ~ p_grid, type = "l")

# (3) L, W, W, L, W, W, W
likelihood_3 <- dbinom(5, size = 7, prob = p_grid)
posterior_3 <- likelihood_3 * prior / sum(likelihood_3 * prior)
plot(posterior_3 ~ p_grid, type = "l")

```

2M2.

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 100)
prior <- ifelse(p_grid < .5, 0, 1) # eliminates the values <.5

# (1) W, W, W
likelihood_1 <- dbinom(3, size = 3, prob = p_grid)
posterior_1 <- likelihood_1 * prior / sum(likelihood_1 * prior)
plot(posterior_1 ~ p_grid, type = "l")

# (2) W, W, W, L
likelihood_2 <- dbinom(3, size = 4, prob = p_grid)
posterior_2 <- likelihood_2 * prior / sum(likelihood_2 * prior)
plot(posterior_2 ~ p_grid, type = "l")

# (3) L, W, W, L, W, W, W
likelihood_3 <- dbinom(5, size = 7, prob = p_grid)
posterior_3 <- likelihood_3 * prior / sum(likelihood_3 * prior)
plot(posterior_3 ~ p_grid, type = "l")
```

2M3.

```{r}
"""
Pr(land|E) = 1 − .7 = .3
Pr(land|M) = 1

Pr(E) = .5
Pr(M) = .5

Pr(E|land) = 
Pr(land|E)*Pr(E) / Pr(land)

Pr(E|land) = 
Pr(land|E)*Pr(E) / Pr(land|E)*Pr(E) + Pr(land|M)*Pr(M)
"""

0.3*0.5 / ( 0.3*0.5 + 1*0.5 )
```

For the exercises below, I highly suggest you to grab a piece of paper and try to solve the problems in a 'visual' way.

2M4.

```{r}

There are three ways to see a black side. 2x from black-black and 1x time from black-white. So 2/3 ways are possible for the other side to be black.

```

2M5.

```{r}

Now there are 5 ways, 4x from 2x black-black and 1x from black-white. The probability is 4/5.

```

2M6.

```{r}

2x b-b 1x b-w 0x w-w #black side
1d b-b 2d b-w 3d w-w #card drawn

2x * 1d = 2 b-b
1x * 2d = 2 b-w

4 ways to see a black card, two of those are black-black so 2/4 = .5 probability

```

2M7.

```{r}
b-b:
  1. card: b-b or b-w (2 ways)
  2. card: b-w or b-b (3 ways)
  2 * 3 = 6
b-w:
  1. card: b-w (1 way)
  2. card: w-w (2 ways)
  1 * 2 = 2
w-w: 0

6 + 2 = 8 ways for black side up, then white side up
so for b-b, 6/8 = .75
```

### Hard

2H1.

```{r}
conditional probability
Pr(twins2|twins1) =
  Pr(twins1, twins2) / Pr(twins)

Pr(twins) = 1/2(.1)+1/2(.2) = .15
Pr(twins1, twins2) = 1/2(.1)+1/2(.4) = .025

Pr(twins2|twins1) = .025/.15 = 1/6
```

2H2.
```{r}
Pr(A|twins1) = 
  Pr(twins1|A)*Pr(A) / Pr(twins) = (.1*.5)/.15 = 1/3
```

2H3.
```{r}
Pr(A|S) =
  Pr(S|A)*Pr(A) / Pr(S)

Pr(S|A) = 1 - .1 = .9; Pr(S|B) = .8
Pr(S) = Pr(S|A)*Pr(A) + Pr(S|B)*Pr(B) = .9*1/3 + .8*2/3 = 5/6

Pr(A|S) = (.9*1/3)/5/6 = .36
```

2H4.
```{r}
Pr(tA|A) = .8
Pr(tB|B) = .65
Pr(tA|B) = 1 - .65 = .35

Pr(A) = .5 # prior

Pr(A|tA) = .8*.5 / .8*.5+.35*.5 = .7
# Pr(tA|A)*Pr(A) / Pr(tA|A)*Pr(A)+Pr(tA|B)*Pr(A)

.5 -> .7 # test validity 

# add birth info as a prior
# Pr(tA|A)*Pr(A) / Pr(tA)
Pr(A) = .36 # prior
Pr(A|tA) = .8*.36 / .36*.8+(1-.36)*.35 = .56 # species A -> test A -> twins (first birth) -> singleton (second birth)
```
